Draft Exploration

### 2020-04-18 draft_exploration.Rmd
added mlb career length to master data

@@

```{r, echo=FALSE}
redraft_table %>%
  ggplot(aes(round, mean_round_diff)) +
  geom_point(na.rm = TRUE) +
  geom_line(aes(group = start_year, color = start_year), na.rm = TRUE) +
  labs(
    x = "Round Range Drafted out of High School",
    y = "Average Improvement Re-draft Round",
    color = "Time Frame",
    title = "Average Change in Round of Re-drafted Unsigned High School Picks",
    subtitle = "Organized by initial round range and time period"
  ) +
  theme_light() +
  scale_color_viridis_c() +
  scale_x_continuous(breaks = c(5, 10, 20, 30, 40, 50)) +
  scale_y_continuous(breaks = c(-10, -5, 0, 10, 20)) +
  geom_hline(color = "red", linetype = "dashed", yintercept = 0)
```
Note estimated intercept with y=0 to see when on average it becomes beneficial
to a player's redraft stock to not sign out of high school. Important to keep in
mind this is for players who are redrafted, so back half of draft improvement
likely partly explained by fewer, the best, initial late round high school picks
being redrafted.

>>>>

# Making table of only mlb_ID and the player's first and final year in the MLB
# to derive MLB career length and join to master.

# NOTE: I kept getting more observations in the new master set after joining
# mlb_career with the left join because guys who played for multiple teams in
# their first or last year were counted that many times in the data so they
# matched multiple times when joined. Therefore, I grouped each set by mlb_ID
# and year and sliced only one observation - it didn't matter which because we
# only need the year.

mlb_start <- war %>%
  filter(!is.na(mlb_ID)) %>%
  group_by(mlb_ID) %>%
  filter(year_ID == min(year_ID)) %>%
  select(mlb_ID, year_ID) %>%
  rename(mlb_start = year_ID) %>%
  group_by(mlb_ID, mlb_start) %>%
  slice(1) %>%
  ungroup()

mlb_end <- war %>%
  filter(!is.na(mlb_ID)) %>%
  group_by(mlb_ID) %>%
  filter(year_ID == max(year_ID)) %>%
  select(mlb_ID, year_ID) %>%
  rename(mlb_end = year_ID) %>%
  group_by(mlb_ID, mlb_end) %>%
  slice(1) %>%
  ungroup()

# joined mlb first year and last year to create career_length by subtracting
# first year from last year in MLB and adding 1 so if career was 1999 only it
# counts as one season.

mlb_career <- mlb_start %>%
  left_join(mlb_end, by = "mlb_ID") %>%
  mutate(mlb_length = mlb_end - mlb_start + 1) %>%
  select(mlb_ID, mlb_length)

# joining mlb career length set to master table

master <- master %>%
  left_join(mlb_career, by = c("mlbid" = "mlb_ID"))

@@

```{r, echo=FALSE}
redraft_table %>%
  ggplot(aes(round, mean_round_diff)) +
  geom_point(na.rm = TRUE) +
  geom_line(aes(group = start_year, color = start_year), na.rm = TRUE) +
  labs(
    x = "Round Range Drafted out of High School",
    y = "Average Improvement Re-draft Round",
    color = "Time Frame",
    title = "Average Change in Round of Re-drafted Unsigned High School Picks",
    subtitle = "Organized by initial round range and time period"
  ) +
  theme_light() +
  scale_color_viridis_c() +
  scale_x_continuous(breaks = c(5, 10, 20, 30, 40, 50)) +
  scale_y_continuous(breaks = c(-10, -5, 0, 10, 20)) +
  geom_hline(color = "red", linetype = "dashed", yintercept = 0)
```

Note estimated intercept with y=0 to see when on average it becomes beneficial
to a player's redraft stock to not sign out of high school. Important to keep in
mind this is for players who are redrafted, so back half of draft improvement
likely partly explained by fewer, the best, initial late round high school picks
being redrafted.
